---
title: "Apply methods controlling FDR to six computational biology case studies"
author: 
- name: Keegan Korthauer
- name: Patrick Kimes
- name: Stephanie Hicks
date: "`r Sys.Date()`"
package: benchmarkfdrData2019
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Apply FDR methods to case studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This vignette describes how each of the methods controlling the 
False Discovery Rate (FDR) was applied to the collection of 
data sets in order to generate the benchmarking result summaries
provided with the package. It also shows how to apply a new FDR 
method to the included data sets, to generate results that can 
be compared to those already included.

## Load data 

```{r}
suppressPackageStartupMessages({
  library(benchmarkfdrData2019)
})

sb <- h3k4me3-promoters-results()
sb 
```

