---
title: "Apply methods controlling FDR to six computational biology case studies"
author: 
- name: Keegan Korthauer
  affiliation: Dana-Farber Cancer Institute
- name: Patrick Kimes
  affiliation: Dana-Farber Cancer Institute
- name: Stephanie Hicks
  affiliation: Johns Hopkins University
output:
  BiocStyle::html_document:
    toc_float: true
package: benchmarkfdrData2019
abstract: |
  Instructions on how to re-create results from Korthauer and Kimes et al. (2019) paper on benchmarking methods that control the false discovery rate (FDR).
vignette: |
  %\VignetteIndexEntry{Apply methods controlling FDR to real case studies and simulated data}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, echo=FALSE, results="hide", message=FALSE}
require(knitr)
opts_chunk$set(error=FALSE, message=FALSE, warning=FALSE)
```

```{r style, echo=FALSE, results='asis'}
BiocStyle::markdown()
```

# Introduction

This package provides a _R_ / _Bioconductor_ resource to describe
how to re-create one of the results from the 
[Korthauer and Kimes et al. (2019)](https://www.biorxiv.org/content/10.1101/458786v1)
which compare methods controlling the False Discovery Rate (FDR). 
In this paper, the methods were applied to the collection of 
data sets in order to generate the benchmarking result summaries
provided with the package. Here, we give an example of how to 
re-create the one of the results. 

## Load data 

We use the `benchmarkfdrData2019` function to download the relevant files
from _Bioconductor_'s ExperimentHub web resource. This includes the
`SummarizedBenchmark` files from the `r Biocpkg("SummarizedBenchmark")` 
package.

```{r}
library(SummarizedBenchmark)
sb <- h3k4me3-promoters-results()
sb 
```

# Session information

```{r}
sessionInfo()
```
